Extracting sunbeams from cucumbers: Computational reproducibility is essential for archaeological science

### 2025-02-10 paper.qmd
writing conclusion

I have presented a bibliometric analysis on the status of archaeology as a science, showing that it is If archaeological science is to continue to progress through increased consensus through the accumulation of reliable facts and methods, it is essential for researchers to take computational reproducibility seriously. Computers have become a central field and laboratory instrument for so much of our work, so we have an ethical duty to document how we change our data as it flows through silicon just as carefully as we document the operating parameters of a mass spectrometer or any other field or laboratory instrument. However, the current state of quantitative archaeology, with most researchers not using open source code, is comparable to the secrecy of alchemy prior to the emergence of chemistry. Abandoning this habit of secrecy in favour of reproducibility is vital if we are to avoid a future where our journals are filled with pretty pictures that the reader has no hope of repeating or adapting in their own work. It is time for computational reproducibility to be considered a minimum requirement for evaluating the integrity and usefulness of empirical results.

>>>>

In Jonathon Swift's classic satirical novel *Gulliver's Travels* (1726) Gulliver visits the fictional Grand Academy of Lagado in Balnibarbi, a caricature of the Royal Society of London, and meets severals researchers working on wildly impractical projects, including one attempting to extracting sunbeams out of cucumbers.

I have presented a bibliometric analysis on the status of archaeology as a science, showing that we generally behave as social scientists If archaeological science is to continue to progress through increased consensus through the accumulation of reliable facts and methods, it is essential for researchers to take computational reproducibility seriously. Computers have become a central field and laboratory instrument for so much of our work, so we have an ethical duty to document how we change our data as it flows through silicon just as carefully as we document the operating parameters of a mass spectrometer or any other field or laboratory instrument. However, the current state of quantitative archaeology, with most researchers not using open source code, is comparable to the secrecy of alchemy prior to the emergence of chemistry. Abandoning this habit of secrecy in favour of reproducibility is vital if we are to avoid a future where our journals are filled with pretty pictures that the reader has no hope of repeating or adapting in their own work. It is time for computational reproducibility to be considered a minimum requirement for evaluating the integrity and usefulness of empirical results.
